---
title: "Econ HW 2"
author: "Maddie Berger & Sara Orofino"
date: "4/25/2019"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, include=FALSE}
# Packages

library(tidyverse)
library(janitor)
library(RColorBrewer)

#Read in data 
demand <- read_csv("Gas_Data.csv")

```


###1. Aggregate Demand, Supply, and Surplus 
  
```{r, include=FALSE}
# Find demand curves using linear regression:

demand_high <- lm(data = demand, price ~ qhigh)
demand_low <- lm(data = demand, price ~ qlow)

int_h <- demand_high$coefficients[1]
q_high <- abs(demand_high$coefficients[2])

int_l <- demand_low$coefficients[1]
q_low <- abs(demand_low$coefficients[2])
```

**High income demand curve:**  
$Price = (`r int_h`) - (`r q_high`)Q$ 

**Low income demand curve:**    
$Price = (`r int_l`) - (`r q_low`)Q$ 

####a. Find aggregate demand   

```{r, include=FALSE}
# Writing functions for demand:

d_high <- function(q) {
  int_h - (q_high*q)
}

d_low <- function(q) {
  int_l - (q_low*q)
}

# Demand as a function of quantity add together for aggregate demand:

d_highq <- function(p){
  (int_h/q_high) - (p/q_high)
}

d_lowq <- function(p) {
  (int_l/q_low) - (p/q_low)
}


# Find the slope and intercept for the aggregate demand:
agg_int <- (int_h/q_high) + (int_l/q_low)
agg_slope <- abs((-1/q_high) + (-1/q_low))

# Write aggregate demand function as function of price:
d_aggq <- function(p) {
  agg_int - agg_slope*p
}

# Change to a function of quantity:
d_agg <- function(q) {
  (agg_int/agg_slope) - ((1/agg_slope)*q)
}

# Quantity at equilibrium price of 5: 
q_sq <- d_highq(5) + d_lowq(5)

# Find mpc by plugging in 5 for the AD curve:

mpc_slope <- (5/q_sq)

mpc <- function(q) {
  mpc_slope*q
}

mec <- function(y) {
  2
}


```

**Aggregate demand curve:**          
$Price = (`r agg_int/agg_slope`) - (`r 1/agg_slope`)Q$   

####b. Find the supply curve  

**Supply curve:**    
$Price = (`r mpc_slope`)*Q$   

```{r, echo=FALSE, message=FALSE, warning=FALSE}
#graphing the functions 

ggplot(data.frame(x=c(70000:200000)), aes(x=x)) + 
  stat_function(fun=d_high, geom ="line", color = "seagreen") + 
  stat_function(fun=d_low, geom = "line", color = "goldenrod") +
  stat_function(fun=mpc, geom = "line") +
  stat_function(fun=mec, geom = "line", color = "red") +
  stat_function(fun=d_agg, geom = "line", color = "blue")+
  labs(x = "Quantity of Gas", y = "Price")+
  scale_x_continuous(limits =c(0,500000), expand = c(0,0))+
  scale_y_continuous(limits=c(0,25), expand=c(0,0))+
  theme_bw()

```


####c. Surplus under the status quo  

```{r, include=FALSE}

######## Consumer Surplus using integration 

# First find quantity where agg demand starts - quantity where price is equal to intercept of d_low
q_cshigh <- d_highq(int_l)

# Integrate to find area under curve for the high demand through quantity where agg starts 
area_high <- integrate(d_high, lower = 0, upper = q_cshigh)

# Integrate agg demand from quantity it starts through the equilibrium quantity 
area_agg <- integrate(d_agg, lower = q_cshigh, upper = q_sq)

# Find total cost at equilibrium (price*quantity)
cost <- 5*q_sq

# Add integrals and subtract cost  
cs_total <- (area_high$value + area_agg$value) - cost


######### Producer Surplus using integration

# Integrate MPC find the area under curve through equilibrium quantity 
mpc_area <- integrate(mpc, lower = 0, upper = q_sq)

# Find PS by subtracting area from cost 
ps_total <- cost - mpc_area$value
```


**Consumers:**    
$CS = (`r cs_total`)$

**Producers:**     
$PS = (`r ps_total`)$  

####d. Environmental Cost under the Status Quo  

```{r, include=FALSE}

tec <- 2*q_sq
  
```


$Environmental~Cost = (`r tec`)$  

###2. Division of Consumer Benefit  

```{r, include=FALSE}
# CS for high demand - area under curve from 0 to quantity at price 5 - (quantity*5)
q_high_sq <- d_highq(5)
fullarea_high <- integrate(d_high, lower = 0, upper = q_high_sq)
cost_high <- q_high_sq * 5
cs_high <- fullarea_high$value - cost_high

# CS for low demand - area under curve from 0 to quantity at price 5 - (quantity*5) 
q_low_sq <- d_lowq(5)
fullarea_low <- integrate(d_low, lower = 0, upper = q_low_sq)
cost_low <- q_low_sq * 5
cs_low <- fullarea_low$value - cost_low
```

**Consumer Surplus for High Income**  
$CS_{High} = (`r cs_high`)$  

**Consumer Surplus for Low Income**  
$CS_{Low} = (`r cs_low`)$  

###3. Implement a Gas Tax of $0.50/gallon  

####a. New quantity of gasoline  

```{r, include=FALSE}

# Suppose the tax is placed on producers and the supply curve shifts left (new intercept 0.50) 

# New equation for MPC with tax:

mpc_tax0.5 <- function(q) {
  0.50 + mpc_slope*q
}

# Find quantity where the MPC and aggregate demand intersect
q_tax0.5 <- ((agg_int/agg_slope)-0.50)/((1/agg_slope)+mpc_slope)
```

$Q_{Tax} = (`r q_tax0.5`)$  

####b. New price of gasoline  

```{r, include=FALSE}
# Price from aggregate demand with the new quantity

p_tax0.5 <- d_agg(q_tax0.5)
```

$P_{Tax} = (`r p_tax0.5`)$  

####c. Surplus to high income consumers  

```{r, include=FALSE}

# CS for high demand = area under curve from 0 to quantity at new price - (new quantity*new price)
q_high_tax0.5 <- d_highq(p_tax0.5)
area_high_tax0.5 <- integrate(d_high, lower = 0, upper = q_high_tax0.5)
cost_high_tax0.5 <- q_high_tax0.5 * p_tax0.5
cs_high_tax0.5 <- area_high_tax0.5$value - cost_high_tax0.5

```

$CS_{High} = (`r cs_high_tax0.5`)$  

####d. Surplus to low income consumers  

```{r, include=FALSE}

# CS for low demand = area under curve from 0 to quantity at new price - (new quantity*new price) 
q_low_tax0.5 <- d_lowq(p_tax0.5)
area_low_tax0.5 <- integrate(d_low, lower = 0, upper = q_low_tax0.5)
cost_low_tax0.5 <- q_low_tax0.5 * p_tax0.5
cs_low_tax0.5 <- area_low_tax0.5$value - cost_low_tax0.5
```

$CS_{Low} = (`r cs_low_tax0.5`)$  

####e. Producer surplus 

```{r, include=FALSE}

# PS = (new quantity*price at MPC) - area under curve through new quantity 
mpc_area_tax0.5 <- integrate(mpc_tax0.5, lower = 0, upper = q_tax0.5)
cost_tax0.5 <- q_tax0.5 * p_tax0.5 
ps_tax0.5 <- cost_tax0.5 - mpc_area_tax0.5$value
```

$PS_{Tax} = (`r ps_tax0.5`)$  

####f. Environmental damage   

```{r, include=FALSE}

tec_tax0.5 <- 2*q_tax0.5
```

$TEC_{Tax} = (`r tec_tax0.5`)$  

####g. Tax revenue 

```{r, include=FALSE}
# Revenue from taxes = (quantity with tax) * (price of the tax)

# Calculate Revenue:
tax0.5_rev <- q_tax0.5 * 0.50

```

$Tax~Revenue = (`r tax0.5_rev`)$  

###4. Tax Revenues for Infrastructure Repairs

```{r, include=FALSE}
# If revenues are being used for infrastructure repairs that area becomes part of CS

# Part that goes to low income consumers is (Qlow)/(total Q with tax) * Rev

########## Generic functions for calculations assuming tax is placed on producers

# Amount of the tax - create a vector that will find tax amounts in increments $0.25 from $0.25-$5
amt_tax <- seq(0.25,5, by = 0.25)
  
#New MPC with the tax (intercept is equal to amount of the tax)
mpc_tax <- function(q) {
  amt_tax + mpc_slope*q
}

# Price at equilibrium (status quo)
p_sq <- 5

#Quantity with tax:
q_tax <- ((agg_int/agg_slope)-amt_tax)/((1/agg_slope)+mpc_slope)

#Price with tax:
p_tax <- d_agg(q_tax)

#Price of MPC with quantity of tax:
p_mpc_tax <- mpc(q_tax)

#Environmental cost of tax:
tec_tax <- 2*q_tax

# Total Revenue
rev_tax <- amt_tax * q_tax

```

####a. Surplus to high income consumers

```{r, include=FALSE}
# CShigh = [(area under curve from 0 to quantity with tax) - (price with tax *quantity with tax)] + Revenue for high income consumers

q_high_tax <- d_highq(p_tax)

# Revenue for High Income Consumers: 
rev_highcon <- rev_tax * ((q_high_tax)/q_tax)

# Define the integration function:

area_h <- function(q) {
  integrate(d_high, lower = 0, upper = q)$value
}

# Use sapply() to integrate to the quantity at each value of the tax:
high_areas <- sapply(q_high_tax, area_h)

# Cost:
cost_high_tax <- q_high_tax * p_tax

# Calculate CS:
cs_high_tax <- (high_areas - cost_high_tax) + rev_highcon
```


####b. Surplus to low income consumers

```{r, include=FALSE}
# CSlow = [(area under curve from 0 to quantity with tax) - (price with tax * quantity with tax)] + Rev for low income consumers - TEC with tax

q_low_tax <- d_lowq(p_tax)

# Revenue for Low Incoming Consumers:
rev_lowcon <- rev_tax * ((q_low_tax)/q_tax)

# Define the integration function:

area_l <- function(q) {
  integrate(d_low, lower = 0, upper = q)$value
}

# Use sapply() to integrate to the quantity at each value of the tax:
low_areas <- sapply(q_low_tax, area_l)

# Cost:
cost_low_tax <- q_low_tax * p_tax

# Calculate CS:
cs_low_tax <- (low_areas - cost_low_tax) + rev_lowcon - tec_tax

```


####c. Surplus to producers  

```{r, include=FALSE}
# PS = [(price with tax * quantity with tax) - area under curve of MPC+tax through quantity with tax]

# Define the integration function:

area_p <- function(q){
  integrate(mpc, lower = 0, upper = q)$value
}

# Use sapply() to integrate to the quantity at each value of the tax:
prod_areas <- sapply(q_tax, area_p)

# Cost using the original supply curve and price at MPC
cost_tax <- q_tax * p_mpc_tax 

#Calculate PS:
ps_tax <- cost_tax - prod_areas


```

```{r, echo=FALSE}

# Bind the baseline to the amount of tax, CS high, CS low, and PS into a dataframe:

tax_scenarios <- as.data.frame(cbind(amt_tax, cs_high_tax, cs_low_tax, ps_tax))


# Graph them:

taxes <- ggplot(tax_scenarios, aes(x = cs_high_tax, y = amt_tax)) +
  geom_line()+
  geom_line(aes(x=cs_low_tax, y = amt_tax)) + 
  geom_line(aes(x=ps_tax, y = amt_tax)) + 
  scale_x_continuous(limits = c(560000,1650000))

# Graph is bad.... try table instead


# Use percent change from the baseline
tax_table <- tax_scenarios %>%
  mutate(cs_high_change = (cs_high_tax-cs_high)/cs_high) %>% 
  mutate(cs_low_change = (cs_low_tax-cs_low)/cs_low) %>% 
  mutate(ps_change = (ps_tax-ps_total)/ps_total) %>% 
  select(amt_tax, cs_low_change, cs_high_change, ps_change)
  

```

