---
title: "Econ HW 2"
author: "Maddie Berger & Sara Orofino"
date: "4/25/2019"
output:
  html_document:
    df_print: paged
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, include=FALSE}
# Packages

library(tidyverse)
library(janitor)
library(RColorBrewer)
library(kableExtra)

#Read in data 
demand <- read_csv("Gas_Data.csv")

```


###1. Aggregate Demand, Supply, and Surplus 
  
```{r, include=FALSE}
# Find demand curves using linear regression:

demand_high <- lm(data = demand, price ~ qhigh)
demand_low <- lm(data = demand, price ~ qlow)

int_h <- demand_high$coefficients[1]
q_high <- abs(demand_high$coefficients[2])

int_l <- demand_low$coefficients[1]
q_low <- abs(demand_low$coefficients[2])
```

**High income demand curve:**  
$Price = `r int_h` - (`r q_high`)Q$ 

**Low income demand curve:**    
$Price = `r int_l` - (`r q_low`)Q$ 

####a. Find aggregate demand   

```{r, include=FALSE}
# Writing functions for demand:

d_high <- function(q) {
  int_h - (q_high*q)
}

d_low <- function(q) {
  int_l - (q_low*q)
}

# Demand as a function of quantity add together for aggregate demand:

d_highq <- function(p){
  (int_h/q_high) - (p/q_high)
}

d_lowq <- function(p) {
  (int_l/q_low) - (p/q_low)
}


# Find the slope and intercept for the aggregate demand:
agg_int <- (int_h/q_high) + (int_l/q_low)
agg_slope <- abs((-1/q_high) + (-1/q_low))

# Write aggregate demand function as function of price:
d_aggq <- function(p) {
  agg_int - agg_slope*p
}

# Change to a function of quantity:
d_agg <- function(q) {
  (agg_int/agg_slope) - ((1/agg_slope)*q)
}

# Quantity at equilibrium price of 5: 
q_sq <- d_highq(5) + d_lowq(5)

# Find mpc by plugging in 5 for the AD curve:

mpc_slope <- (5/q_sq)

mpc <- function(q) {
  mpc_slope*q
}

mec <- function(y) {
  2
}


```

**Aggregate demand curve:**          
$Price = `r agg_int/agg_slope` - (`r 1/agg_slope`)Q$   

####b. Find the supply curve  

**Supply curve:**    
$Price = (`r mpc_slope`)*Q$   

```{r, echo=FALSE, message=FALSE, warning=FALSE}
#graphing the functions 

ggplot(data.frame(x=c(70000:200000)), aes(x=x)) + 
  stat_function(fun=d_high, geom ="line", color = "seagreen") + 
  stat_function(fun=d_low, geom = "line", color = "goldenrod") +
  stat_function(fun=mpc, geom = "line") +
  stat_function(fun=mec, geom = "line", color = "red") +
  stat_function(fun=d_agg, geom = "line", color = "blue")+
  labs(x = "Quantity of Gas", y = "Price")+
  scale_x_continuous(limits =c(0,500000), expand = c(0,0))+
  scale_y_continuous(limits=c(0,25), expand=c(0,0))+
  theme_bw()

```


####c. Surplus under the status quo  

```{r, include=FALSE}

######## Consumer Surplus using integration 

# First find quantity where agg demand starts - quantity where price is equal to intercept of d_low
q_cshigh <- d_highq(int_l)

# Integrate to find area under curve for the high demand through quantity where agg starts 
area_high <- integrate(d_high, lower = 0, upper = q_cshigh)

# Integrate agg demand from quantity it starts through the equilibrium quantity 
area_agg <- integrate(d_agg, lower = q_cshigh, upper = q_sq)

# Find total cost at equilibrium (price*quantity)
cost <- 5*q_sq

# Add integrals and subtract cost  
cs_total <- ((area_high$value + area_agg$value) - cost)/1000000


######### Producer Surplus using integration

# Integrate MPC find the area under curve through equilibrium quantity 
mpc_area <- integrate(mpc, lower = 0, upper = q_sq)

# Find PS by subtracting area from cost 
ps_total <- cost - mpc_area$value
ps_total_mil <- ps_total/1000000
```


**Consumers:**    
$CS = `r cs_total`~Million~USD$ 

**Producers:**     
$PS = `r ps_total_mil`~Million~USD$ 

####d. Environmental Cost under the Status Quo  

```{r, include=FALSE}

tec <- (2*q_sq)/1000000
  
```


$Environmental~Cost = `r tec`~Million~USD$ 

###2. Division of Consumer Benefit  

```{r, include=FALSE}
# CS for high demand - area under curve from 0 to quantity at price 5 - (quantity*5)
q_high_sq <- d_highq(5)
fullarea_high <- integrate(d_high, lower = 0, upper = q_high_sq)
cost_high <- q_high_sq * 5
cs_high <- fullarea_high$value - cost_high
cs_high_mil <- cs_high/1000000

# CS for low demand - area under curve from 0 to quantity at price 5 - (quantity*5) 
q_low_sq <- d_lowq(5)
fullarea_low <- integrate(d_low, lower = 0, upper = q_low_sq)
cost_low <- q_low_sq * 5
cs_low <- fullarea_low$value - cost_low
cs_low_mil <- cs_low/1000000
```

**Consumer Surplus for High Income**  
$CS_{High} = `r cs_high_mil`~Million~USD$ 

**Consumer Surplus for Low Income**  
$CS_{Low} = `r cs_low_mil`~Million~USD$ 

###3. Implement a Gas Tax of $0.50/gallon  

####a. New quantity of gasoline  

```{r, include=FALSE}

# Suppose the tax is placed on producers and the supply curve shifts left (new intercept 0.50) 

# New equation for MPC with tax:

mpc_tax0.5 <- function(q) {
  0.50 + mpc_slope*q
}

# Find quantity where the MPC and aggregate demand intersect
q_tax0.5 <- ((agg_int/agg_slope)-0.50)/((1/agg_slope)+mpc_slope)
```

$Q_{Tax} = `r q_tax0.5`$  

####b. New price of gasoline  

```{r, include=FALSE}
# Price from aggregate demand with the new quantity

p_tax0.5 <- d_agg(q_tax0.5)
```

$P_{Tax} = `r p_tax0.5`$  

####c. Surplus to high income consumers  

```{r, include=FALSE}

# CS for high demand = area under curve from 0 to quantity at new price - (new quantity*new price)
q_high_tax0.5 <- d_highq(p_tax0.5)
area_high_tax0.5 <- integrate(d_high, lower = 0, upper = q_high_tax0.5)
cost_high_tax0.5 <- q_high_tax0.5 * p_tax0.5
cs_high_tax0.5 <- (area_high_tax0.5$value - cost_high_tax0.5)/1000000

```

$CS_{High} = `r cs_high_tax0.5`~Million~USD$   

####d. Surplus to low income consumers  

```{r, include=FALSE}

# CS for low demand = area under curve from 0 to quantity at new price - (new quantity*new price) 
q_low_tax0.5 <- d_lowq(p_tax0.5)
area_low_tax0.5 <- integrate(d_low, lower = 0, upper = q_low_tax0.5)
cost_low_tax0.5 <- q_low_tax0.5 * p_tax0.5
cs_low_tax0.5 <- (area_low_tax0.5$value - cost_low_tax0.5)/1000000
```

$CS_{Low} = `r cs_low_tax0.5`~Million~USD$     

####e. Producer surplus 

```{r, include=FALSE}

# PS = (new quantity*price at MPC) - area under curve through new quantity 
mpc_area_tax0.5 <- integrate(mpc_tax0.5, lower = 0, upper = q_tax0.5)
cost_tax0.5 <- q_tax0.5 * p_tax0.5 
ps_tax0.5 <- (cost_tax0.5 - mpc_area_tax0.5$value)/1000000
```

$PS_{Tax} = `r ps_tax0.5`~Million~USD$   

####f. Environmental damage   

```{r, include=FALSE}

tec_tax0.5 <- (2*q_tax0.5)/1000000
```

$TEC_{Tax} = `r tec_tax0.5`~Million~USD$    

####g. Tax revenue 

```{r, include=FALSE}
# Revenue from taxes = (quantity with tax) * (price of the tax)

# Calculate Revenue:
tax0.5_rev <- (q_tax0.5 * 0.50)/1000000

```

$Tax~Revenue = `r tax0.5_rev`~Million~USD$     

###4. Tax Revenues for Infrastructure Repairs

```{r, include=FALSE}
# If revenues are being used for infrastructure repairs that area becomes part of CS

# Part that goes to low income consumers is (Qlow)/(total Q with tax) * Rev

########## Generic functions for calculations assuming tax is placed on producers

# Amount of the tax - create a vector that will find tax amounts in increments $0.25 from $0.25-$5
amt_tax <- seq(0.25,5, by = 0.25)
  
#New MPC with the tax (intercept is equal to amount of the tax)
mpc_tax <- function(q) {
  amt_tax + mpc_slope*q
}

# Price at equilibrium (status quo)
p_sq <- 5

#Quantity with tax:
q_tax <- ((agg_int/agg_slope)-amt_tax)/((1/agg_slope)+mpc_slope)

#Price with tax:
p_tax <- d_agg(q_tax)

#Price of MPC with quantity of tax:
p_mpc_tax <- mpc(q_tax)

#Environmental cost of tax:
tec_tax <- 2*q_tax

# Total Revenue
rev_tax <- amt_tax * q_tax

```

####a. Surplus to high income consumers

```{r, echo=FALSE}
# CShigh = [(area under curve from 0 to quantity with tax) - (price with tax *quantity with tax)] + Revenue for high income consumers

q_high_tax <- d_highq(p_tax)

# Revenue for High Income Consumers: 
rev_highcon <- rev_tax * ((q_high_tax)/q_tax)

# Define the integration function:

area_h <- function(q) {
  integrate(d_high, lower = 0, upper = q)$value
}

# Use sapply() to integrate to the quantity at each value of the tax:
high_areas <- sapply(q_high_tax, area_h)

# Cost:
cost_high_tax <- q_high_tax * p_tax

# Calculate CS:
cs_high_tax <- (high_areas - cost_high_tax) + rev_highcon

#Output information in table:
cs_high_table <- as.data.frame(cbind(amt_tax, cs_high_tax)) %>% 
  mutate(cs_high_mil = cs_high_tax/1000000) %>% 
  select(amt_tax, cs_high_mil)

kable(cs_high_table, col.names = c("Tax Amount", "High Income Consumer Welfare (Million USD)"), caption = "Table 1. High Income Consumer Welfare at Variable Gas Tax Amounts", digits = 4) %>% 
  kable_styling(bootstrap_options = c("striped", "condensed", "bordered"), full_width = FALSE)

```


####b. Surplus to low income consumers

```{r, echo=FALSE}
# CSlow = [(area under curve from 0 to quantity with tax) - (price with tax * quantity with tax)] + Rev for low income consumers - TEC with tax

q_low_tax <- d_lowq(p_tax)

# Revenue for Low Incoming Consumers:
rev_lowcon <- rev_tax * ((q_low_tax)/q_tax)

# Define the integration function:

area_l <- function(q) {
  integrate(d_low, lower = 0, upper = q)$value
}

# Use sapply() to integrate to the quantity at each value of the tax:
low_areas <- sapply(q_low_tax, area_l)

# Cost:
cost_low_tax <- q_low_tax * p_tax

# Calculate CS:
cs_low_tax <- (low_areas - cost_low_tax) + rev_lowcon - tec_tax

#Output information in table:
cs_low_table <- as.data.frame(cbind(amt_tax, cs_low_tax)) %>% 
  mutate(cs_low_mil = cs_low_tax/1000000) %>% 
  select(amt_tax, cs_low_mil)

kable(cs_low_table, col.names = c("Tax Amount", "Low Income Consumer Welfare (Million USD)"), caption = "Table 2. Low Income Consumer Welfare at Variable Gas Tax Amounts", digits = 4) %>% 
  kable_styling(bootstrap_options = c("striped", "condensed", "bordered"), full_width = FALSE)

```


####c. Surplus to producers  

```{r, echo=FALSE}
# PS = [(price with tax * quantity with tax) - area under curve of MPC+tax through quantity with tax]

# Define the integration function:

area_p <- function(q){
  integrate(mpc, lower = 0, upper = q)$value
}

# Use sapply() to integrate to the quantity at each value of the tax:
prod_areas <- sapply(q_tax, area_p)

# Cost using the original supply curve and price at MPC
cost_tax <- q_tax * p_mpc_tax 

#Calculate PS:
ps_tax <- cost_tax - prod_areas

#Output information in table:
ps_table <- as.data.frame(cbind(amt_tax, ps_tax)) %>% 
  mutate(ps_mil = ps_tax/1000000) %>% 
  select(amt_tax, ps_mil)

kable(ps_table, col.names = c("Tax Amount", "Producer Welfare (Million USD)"), caption = "Table 3. Producer Welfare at Variable Gas Tax Amounts", digits = 4) %>% 
  kable_styling(bootstrap_options = c("striped", "condensed", "bordered"), full_width = FALSE)


```


####Possible Revenues and Welfare Changes from Gas Tax:  
**Assumptions**  
 - Benefits from infrastructure repairs are proportional to amount driven  
 - Low income consumers pay the entire environmental cost   
 
```{r, echo=FALSE}

# Bind the baseline to the amount of tax, CS high, CS low, and PS into a dataframe:

tax_scenarios <- as.data.frame(cbind(amt_tax, cs_high_tax, cs_low_tax, ps_tax, rev_tax))


# Graph them:

taxes <- ggplot(tax_scenarios, aes(x = cs_high_tax, y = amt_tax)) +
  geom_line()+
  geom_line(aes(x=cs_low_tax, y = amt_tax)) + 
  geom_line(aes(x=ps_tax, y = amt_tax)) + 
  scale_x_continuous(limits = c(560000,1650000))

# Graph is bad.... try table instead


# Use percent change from the baseline
tax_table <- tax_scenarios %>%
  mutate(cs_high_change = ((cs_high_tax-cs_high)/cs_high)*100) %>% 
  mutate(cs_low_change = ((cs_low_tax-cs_low)/cs_low)*100) %>% 
  mutate(ps_change = ((ps_tax-ps_total)/ps_total)*100) %>% 
  mutate(rev_taxmil = rev_tax/1000000) %>% 
  select(amt_tax, rev_taxmil,cs_low_change, cs_high_change, ps_change)
  

taxtable_final <- kable(tax_table, col.names = c("Tax Amount (USD)", "Tax Revenue (Million USD)", "Welfare Change Low Income (%)", "Welfare Change High Income (%)", "Welfare Change Producers (%)"), caption = "Figure 4. Comparison of Revenue and Welfare Potential with Variable Gas Tax Amounts", digits = 4) %>% 
  kable_styling(bootstrap_options = c("striped", "bordered", "condensed"), full_width = F)

taxtable_final


```


###5. Electric cars lower demand for each group by one half (vertically)
```{r, include=FALSE}
######generic functions used for this problem 

int_h_2 <- int_h/2 #divide intercepts by 2 to shift demand curves vertically by 1/2

int_l_2 <- int_l/2

#new demand functions with new intercepts 
d_high_2 <- function(q) {
  int_h_2 - (q_high*q)
}

d_low_2 <- function(q) {
  int_l_2 - (q_low*q)
}

#change to a function of p 

d_highq_2 <- function(p){
  (int_h_2/q_high) - (p/q_high)
}

d_lowq_2 <- function(p) {
  (int_l_2/q_low) - (p/q_low)
}

#new aggregate demand curve 

agg_int_ev <- (int_h_2/q_high) + (int_l_2/q_low)
agg_slope_ev <- abs((-1/q_high) + (-1/q_low))

d_agg_2q <- function(p) {
  agg_int_ev + agg_slope_ev*p
}

d_agg_2p <- function(q) {
  (agg_int_ev/agg_slope_ev) - ((1/agg_slope_ev)*q)
}

```

####a. & b. Gas consumption by High/Low income Consumers
```{r, include=FALSE}
#set equal to MPC curve to find new quantities 

q_ev_low <- int_l_2/(mpc_slope+q_low)

q_ev_high <- int_h_2/(mpc_slope+q_high)

#add quantities together

q_ev_agg <- q_ev_high + q_ev_high
```

####c. New price of gasoline with higher EV use and lower gas demand

```{r, include=FALSE}
#plug agg price into demand agg function to find the price 

d_agg_2p(q_ev_agg)

#double check that this makes sense

d_high_2(q_ev_high)
d_low_2(q_ev_low)
```

####d. Environmental Cost 

```{r, include=FALSE}
env_cost_ev <- (q_ev_agg*2)/1000000
```

###6. Compare a 2.00 per gal tax to the influence of EV

```{r, echo=FALSE}
# slope of MSC with a 2.00 tax
mpc_tax2.0 <- function(q) {
  2.0 + mpc_slope*q
}

# Find quantity where the new MSC and aggregate demand intersect
q_tax2.0 <- ((agg_int/agg_slope)-2.0)/((1/agg_slope)+mpc_slope)

#environmental damage

env_cost_tax2 <- (q_tax2.0*2)/1000000

## put all the environmental damages to compare 

env_damage_table <- as.data.frame(rbind(tec,env_cost_tax2, env_cost_ev)) 

tax2_reduction = ((tec-env_cost_tax2)/tec)*100

ev_reduction = ((tec-env_cost_ev)/tec)*100

env_damage_table$Reduction <- c(0,tax2_reduction, ev_reduction)
  
  
colnames(env_damage_table) <- c("Environmental Cost (USD)", "Reduction (%)")
rownames(env_damage_table) <- c("Baseline", "$2.00 Gas Tax", "High EV Demand")

#make it nice

env_damage_final <- kable(env_damage_table, digits = 4) %>% 
  kable_styling(bootstrap_options = c("striped", "bordered", "condensed"), full_width = F)

env_damage_final
```

